Manuscript figures

### 2025-09-02 manuscriptFigures.Rmd
edits for fig 5-7

@@

@@

>>>>

# params area weighted (only for lagSurfaceRunoff, a_rain, a_snow)
params <- read_csv('./38HRU_dec2024_updates/params_weighted_dec2024.csv') %>%
  filter(param %in% c('lagSurfaceRunoff',
                      'a_rain',
                      'a_snow')) %>% # select params in the top 8 sens that are area weighted
  rename(value = area_weighted_average,
         input_data = source) %>%
  dplyr::select(param, input_data, year_missing, value)

# params unburned (above is params that needed to be area weighted bc burn/unburn)
params_unburn <- read_csv('./38HRU_dec2024_updates/params_unburn.csv') %>%
  filter(param %in% c('soilOutLPS',
                      'snow_trs',
                      'lagInterflow',
                      'baseTemp',
                      't_factor',
                      'soilLatVertLPS')) %>%
  dplyr::select(param, input_data, year_missing, value) %>%
  mutate(year_missing = if_else(year_missing == 'none',
                                'all',
                                year_missing))

# join the two to get the top 8 sensitve params
param_sens <- bind_rows(params, params_unburn) %>% 
  mutate(input_data = if_else(input_data == 'grid',
                              'Gridded',
                              'Station'))

# read in pbias categories
pbias <- read_csv('./38HRU_dec2024_updates/pbiasCats.csv') %>%
  filter(year == c('All')) %>%
  dplyr::select(input_data, year_missing, pbias)

nse <- read_csv('./38HRU_dec2024_updates/nseCats.csv') %>%
  filter(year == c('All')) %>%
  dplyr::select(input_data, year_missing, nse)

perf <- left_join(nse, pbias, by = c('input_data',
                                     'year_missing'))

perf <- perf %>%
  mutate(performance = case_when(
                                nse < 0.5 & abs(pbias) >= 25 ~ 'Neither',
                                nse > 0.5 & abs(pbias) >= 25 ~ 'NSE',
                                nse < 0.5 & abs(pbias) < 25 ~ '% bias satisfactory',
                                nse > 0.5 & abs(pbias) < 25 ~ 'NSE + % bias satisfactory')) %>%
  mutate(input_data = if_else(input_data == 'grid',
                              'Gridded', 'Station'),
         year_missing = if_else(year_missing == 'None',
                                'all', year_missing))

param_sens <- left_join(param_sens,
                        perf,
                        by = c('input_data',
                               'year_missing'))

param_sens$param <- fct_recode(param_sens$param,
  "snow_trs"        = "snow_trs",
  "baseTemp (°C)"      = "baseTemp",
  "t_factor"    = "t_factor",
  "a_rain (mm)"      = "a_rain",
  "a_snow (mm)"      = "a_snow",
  "lagSurfaceRunoff"    = "lagSurfaceRunoff",
  "soilLatVertLPS"      = "soilLatVertLPS",
  "soilOutLPS"       = "soilOutLPS",
  "lagInterflow"         = "lagInterflow"
)

param_sens$param <- fct_relevel(param_sens$param, 
                                "snow_trs",
                                "baseTemp (°C)",
                                "t_factor",
                                "a_rain (mm)",
                                "a_snow (mm)",
                                "lagSurfaceRunoff",
                                "soilLatVertLPS",
                                "soilOutLPS",
                                "lagInterflow")

# boxplots of top 9
ggplot(param_sens, aes(x=input_data, y=value, fill=input_data)) +
  geom_boxplot(alpha=0.8,
                outlier.shape=NA) +
  facet_wrap(~param, scales = 'free_y',
             strip.position = 'left') +
  labs(y=element_blank(),
       x=element_blank()) +
  theme_bw(base_size=20) +
  stat_compare_means(aes(group = input_data),
                     method = "wilcox.test", 
                     label = "p.signif",
                     hide.ns = T,
                     paired = F,
                     #label.y = 0,
                     #label.x = 1.4,
                     size = 7) +
  geom_jitter(aes(shape = performance),
            position = position_jitter(width = 0.1), 
            size = 2.5, 
            alpha = 0.6,
            stroke = 1,
            color = "black") +
  scale_shape_manual(values = c(
  "% bias satisfactory" = 1,
  "NSE + % bias satisfactory" = 16
)) +
  scale_fill_manual(values = c("#0047AB", "orange"),
                    guide='none') +
    theme(
    legend.position = 'bottom',
    #legend.justification = c(0, 1.1), 
    legend.text = element_text(size = 16),
    legend.title = element_blank(), 
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    strip.placement = 'outside',
    #strip.background = element_blank(),
    axis.title.y = element_blank()) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.25)))

ggsave('./figures/figure5.png',
       dpi=600,
       width=12,
       height=8,
       bg='white')

## same as above but with param limits added
param_limits <- data.frame(
  param = c("snow_trs", "baseTemp (°C)", "t_factor",
            "a_rain (mm)", "a_snow (mm)", "lagSurfaceRunoff",
            "soilLatVertLPS", "soilOutLPS", "lagInterflow"),
  low  = c(-5, -5, 0, 0.1, 0.1, 1, 0.001, 0.01, 1),
  high = c(5, 5, 5, 10, 10, 1.5, 10, 10, 2)
)

param_lines <- param_limits %>%
  pivot_longer(cols = c(low, high),
               names_to = "type", 
               values_to = "y")

# Make sure param in param_lines is a factor with the same order
param_lines$param <- fct_relevel(param_lines$param,
                                 "snow_trs",
                                 "baseTemp (°C)",
                                 "t_factor",
                                 "a_rain (mm)",
                                 "a_snow (mm)",
                                 "lagSurfaceRunoff",
                                 "soilLatVertLPS",
                                 "soilOutLPS",
                                 "lagInterflow")

ggplot(param_sens, aes(x=input_data, y=value, fill=input_data)) +
  # add horizontal lines for low/high values
  geom_hline(data = param_lines, aes(yintercept = y),
             linetype = "dashed", color = "grey70") +
  geom_boxplot(alpha=0.8,
                outlier.shape=NA) +
  facet_wrap(~param, scales = 'free_y',
             strip.position = 'left') +
  labs(y=element_blank(),
       x=element_blank()) +
  theme_bw(base_size=20) +
  stat_compare_means(aes(group = input_data),
                     method = "wilcox.test", 
                     label = "p.signif",
                     hide.ns = T,
                     paired = F,
                     #label.y = 0,
                     #label.x = 1.4,
                     size = 7) +
  geom_jitter(aes(shape = performance),
            position = position_jitter(width = 0.1), 
            size = 2.5, 
            alpha = 0.6,
            stroke = 1,
            color = "black") +
  scale_shape_manual(values = c(
  "% bias satisfactory" = 1,
  "NSE + % bias satisfactory" = 16
)) +
  scale_fill_manual(values = c("#0047AB", "orange"),
                    guide='none') +
    theme(
    legend.position = 'bottom',
    #legend.justification = c(0, 1.1), 
    legend.text = element_text(size = 16),
    legend.title = element_blank(), 
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    strip.placement = 'outside',
    #strip.background = element_blank(),
    axis.title.y = element_blank()) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.25)))

ggsave('./figures/figure5_wlimits.png',
       dpi=600,
       width=12,
       height=8,
       bg='white')

### For Supp

# show rest of params for supplemental
# first take area weighted params (except lagSurfaceRunoff, a_rain, a_snow since that is already shown in main fig)
otherParamsWeighted <- read_csv('./38HRU_dec2024_updates/params_weighted_dec2024.csv') %>%
  filter(!param %in% c('lagSurfaceRunoff',
                       'a_rain',
                       'a_snow')) %>%
  rename(value = area_weighted_average,
         input_data = source) %>%
  dplyr::select(param, input_data, value)

# params that were not affected by burn
otherParams <- read_csv('./38HRU_dec2024_updates/params_unburn.csv') %>%
  filter(!param %in% c('soilOutLPS',
                       'snow_trs',
                       'lagInterflow',
                       'baseTemp',
                       't_factor',
                       'lagSurfaceRunoff',
                       'a_rain',
                       'a_snow',
                       'soilLatVertLPS')) %>% # these are the ones already shown
  filter(!param %in% otherParamsWeighted$param) %>% # there are the ones that need to be area weighted
  dplyr::select(param, input_data, value)

# now join the area weighted other to the other params
otherParams <- bind_rows(otherParams, otherParamsWeighted)

# plot the remaning params for catchment
ggplot(otherParams, aes(x=input_data, y=value, fill=input_data)) +
  geom_boxplot(alpha=0.8,
                outlier.colour = "black") +
  facet_wrap(~param, scales = 'free_y',
             ncol=3) +
  labs(y=element_blank(),
       x=element_blank()) +
  theme_bw(base_size=20) +
  stat_compare_means(aes(group = input_data),
                     method = "wilcox.test", 
                     label = "p.signif",
                     hide.ns = T,
                     paired = F,
                     #label.y = 0,
                     #label.x = 1.4,
                     size = 7) +
  scale_fill_manual(values = c("#0047AB", "orange")) +
    theme(
    legend.position = 'none',
    #legend.justification = c(0, 1.1), 
    legend.text = element_text(size = 16),
    legend.title = element_blank(), 
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    #strip.text = element_blank()
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.001, 0.3)))

ggsave('./figures/remainingparams_boxplots_SUPP.png',
       dpi=600,
       height=16,
       width=10)

@@

gw_gridall <- read_csv(here('38HRU_dec2024_updates/catchment_output/grid_all.csv'), skip = 3) %>%
  slice(-c(1,2)) %>%
  dplyr::select(time,
                catchmentRD1_w, 
                catchmentRD2_w, 
                catchmentRG1_w, 
                catchmentRG2_w) %>%
  mutate(source = 'grid_all') %>%
  mutate(across(c(catchmentRD1_w, 
                  catchmentRD2_w, 
                  catchmentRG1_w, 
                  catchmentRG2_w), 
                as.numeric))

gw_statnall <- read_csv(here('38HRU_dec2024_updates/catchment_output/statn_all.csv'), skip = 3) %>%
  slice(-c(1,2)) %>%
  dplyr::select(time, 
                catchmentRD1_w, 
                catchmentRD2_w, 
                catchmentRG1_w, 
                catchmentRG2_w) %>%
  mutate(source = 'statn_all') %>%
  mutate(across(c(catchmentRD1_w, 
                  catchmentRD2_w, 
                  catchmentRG1_w, 
                  catchmentRG2_w), 
                as.numeric))

# combine the two above
gw <- bind_rows(
  gw_gridall, gw_statnall) %>%
  mutate(date = mdy(time),
         year = year(date)) %>%
  arrange(date) %>%
  group_by(year, source) %>%
  mutate(RD1cumsum = cumsum(catchmentRD1_w),
         RD2cumsum = cumsum(catchmentRD2_w),
         RG1cumsum = cumsum(catchmentRG1_w),
         RG2cumsum = cumsum(catchmentRG2_w))

gw_maxcumsum <- gw %>%
  group_by(source, year) %>%
  summarize(yr_cum_RD1 = max(RD1cumsum),
            yr_cum_RD2 = max(RD2cumsum),
            yr_cum_RG1 = max(RG1cumsum),
            yr_cum_RG2 = max(RG2cumsum)) %>%
  mutate(year = as.factor(year)) %>%
  pivot_longer(!c(year, source), names_to = 'flow_type', values_to = 'flow_cms') %>%
  mutate(source = if_else(source == 'grid_all',
                          'Gridded',
                          'Station'))

# make a vector of the letters - this is from the ANOVA+Tukey results in fourmile_output_dec2024
letters_vector <- c("a", "b", "b", "c", "c", "c", "c", "c")

gw_maxcumsum <- gw_maxcumsum %>%
  mutate(letter = case_when(
    source == "Gridded" & flow_type == "yr_cum_RG1" ~ letters_vector[1],
    source == "Station" & flow_type == "yr_cum_RG1" ~ letters_vector[2],
    source == "Station" & flow_type == "yr_cum_RD2" ~ letters_vector[3],
    source == "Gridded" & flow_type == "yr_cum_RD1" ~ letters_vector[4],
    source == "Station" & flow_type == "yr_cum_RD1" ~ letters_vector[5],
    source == "Gridded" & flow_type == "yr_cum_RD2" ~ letters_vector[6],
    source == "Gridded" & flow_type == "yr_cum_RG2" ~ letters_vector[7],
    source == "Station" & flow_type == "yr_cum_RG2" ~ letters_vector[8]
  ))

# now make the same figure as above but with letters
facet_labels <- c('yr_cum_RD1' = 'Surface runoff',
                  'yr_cum_RD2' = 'Interflow',
                  'yr_cum_RG1' = 'Fast groundwater',
                  'yr_cum_RG2' = 'Slow groundwater')

perf <- read_csv('./38HRU_dec2024_updates/nse_pbias_grid_statn_all.csv') %>%
  mutate(year = as.factor(year))

gw_maxcumsum <- left_join(gw_maxcumsum,
                          perf,
                          by = c('source',
                                 'year'))

gw_maxcumsum <- gw_maxcumsum %>%
  mutate(flow_cms = if_else(flow_type == 'yr_cum_RD1' &
                                flow_cms > 48,
                              NA,
                              flow_cms))

gw_maxcumsum_letters <- gw_maxcumsum %>%
  group_by(source, flow_type) %>%
  mutate(max = max(flow_cms)) %>%
  distinct(flow_type, source, .keep_all = TRUE) %>%
    mutate(max = if_else((flow_type == 'yr_cum_RD1' & 
                           source == 'Gridded'),
                         3.5, max)) %>%
    mutate(max = if_else((flow_type == 'yr_cum_RD1' & 
                           source == 'Station'),
                         1.3, max)) %>%
  mutate(y_text = max * 1.06)

flow_box <- ggplot(gw_maxcumsum, aes(x = source, y = flow_cms, fill = source)) +
  geom_boxplot(outliers=F,width=0.7) + # need to note 1 outlier is removed from RD1 grid and statn (2013 flood) ****
  geom_jitter(
    aes(shape = both),
    position = position_jitter(width = 0.25),
    size = 2.5,
    alpha = 0.7,
    stroke = 1,
    color = "black") +
  scale_shape_manual(values = c(
  "NSE + % bias unsatisfactory" = 4,
  "NSE + % bias satisfactory" = 16,
  "NSE satisfactory" = 1)) +
  geom_text(data = gw_maxcumsum_letters, 
            aes(label=letter,
                y=y_text),
            position = position_dodge(width = 0.97),
            size = 6, 
            vjust = 0) + 
  facet_wrap(~flow_type, 
             scales = 'free',
             labeller = labeller(flow_type = facet_labels)) +
  labs(y = 'mm of water') + 
  theme_bw(base_size=20) +
  scale_fill_manual(values = c("#0047AB", "orange"), guide = 'none') +
  theme(
    #legend.position = c(0.94, 0.36),
    legend.position = 'bottom',
    legend.text = element_text(size = 16),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    axis.title.x = element_blank(),
    legend.spacing = unit(0.1, "lines"),
    legend.margin = margin(t = -8, r = 0, b = -2, l = 0)) +
    #axis.text.x = element_blank(),
    #axis.ticks.x = element_blank()) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.15)))

flow_box

ggsave('./figures/figure6.png',
       dpi=600,
       width = 12,
       height = 8)

# get some stats for manuscript
stats <- gw %>%
  group_by(source, year) %>%
  summarize(yr_cum_RD1 = max(RD1cumsum),
            yr_cum_RD2 = max(RD2cumsum),
            yr_cum_RG1 = max(RG1cumsum),
            yr_cum_RG2 = max(RG2cumsum)) %>%
  mutate(year = as.factor(year)) %>%
  pivot_longer(!c(year, source), names_to = 'flow_type', values_to = 'flow_cms') %>%
  mutate(source = if_else(source == 'grid_all',
                          'Gridded',
                          'Station')) %>% 
  group_by(source, year) %>%
  mutate(totFlow = sum(flow_cms)) %>%
  group_by(source, year) %>%
  mutate(fracFlow = (flow_cms/totFlow)*100) #%>%
  group_by(source, flow_type) %>% 
  summarize(meanPercFlow = mean(fracFlow))

@@

statn_hi <- read_csv('./38HRU_dec2024_updates/HRU/statn_HRU_18.csv') %>% 
  mutate(date = mdy(time),
         source = 'statn') %>%
  dplyr::select(-time)

grid_hi <- read_csv('./38HRU_dec2024_updates/HRU/grid_HRU_18.csv') %>%
  mutate(date = mdy(time),
         source = 'grid') %>%
  dplyr::select(-time)

both_hi <- bind_rows(statn_hi, grid_hi) %>%
  mutate(year = year(date),
         month = month(date)) %>%
  filter(year == 2016,
         month %in% c(4,5,6))

statn_low <- read_csv('./38HRU_dec2024_updates/HRU/statn_HRU_19.csv') %>% 
  mutate(date = mdy(time),
         source = 'statn') %>%
  dplyr::select(-time)

grid_low <- read_csv('./38HRU_dec2024_updates/HRU/grid_HRU_19.csv') %>%
  mutate(date = mdy(time),
         source = 'grid') %>%
  dplyr::select(-time)

both_low <- bind_rows(statn_low, grid_low) %>%
  mutate(year = year(date),
         month = month(date)) %>%
  filter(year == 2016,
         month %in% c(4,5,6))

## combine low and high together
hi <- both_hi %>%
  dplyr::select(tmean, snowTotSWE, precip, source, date) %>%
  mutate(elevation = 'High')

low <- both_low %>%
  dplyr::select(tmean, snowTotSWE, precip, source, date) %>%
  mutate(elevation = 'Low')

both <- bind_rows(hi, low) %>%
  mutate(elev_source =
           case_when(source == 'statn' & elevation == 'High' ~ 'Station - High',
                     source == 'statn' & elevation == 'Low' ~ 'Station - Low',
                     source == 'grid' & elevation == 'High' ~ 'Gridded - High',
                     source == 'grid' & elevation == 'Low' ~ 'Gridded - Low'))

# figure
tmean <- ggplot(both,
       aes(x=date, y = tmean, 
           color = elev_source,
           linetype=elev_source)) +
   geom_line(size=1, alpha=0.8) +  
    scale_color_manual(values = c("Gridded - High" = "#0047AB", 
                                  "Gridded - Low" = "#0047AB", 
                                  "Station - High" = "orange",
                                  "Station - Low" = "orange")) +
  scale_linetype_manual(values = c("Gridded - High" = "dashed", 
                                  "Gridded - Low" = "solid", 
                                  "Station - High" = "dashed",
                                  "Station - Low" = "solid")) + 
  scale_x_date(date_labels = "%b") +
  labs(x = element_blank(),
       y = 'Mean temperature (°C)') +
  theme_bw(base_size = 20) +
  theme(legend.position = 'bottom',
    #legend.position = c(0.24, 0.852), 
    legend.text = element_text(size = 10),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
    #plot.margin = margin(t = 10, r = 20, b = 10, l = 10)
  )

swe <- ggplot(both,
       aes(x=date, y = snowTotSWE, 
           color = elev_source,
           linetype=elev_source)) +
   geom_line(size=1, alpha=0.8) +  
    scale_color_manual(values = c("Gridded - High" = "#0047AB", 
                                  "Gridded - Low" = "#0047AB", 
                                  "Station - High" = "orange",
                                  "Station - Low" = "orange")) +
  scale_linetype_manual(values = c("Gridded - High" = "dashed", 
                                  "Gridded - Low" = "solid", 
                                  "Station - High" = "dashed",
                                  "Station - Low" = "solid")) + 
  scale_x_date(date_labels = "%b") +
  #scale_y_log10() +
  labs(x = element_blank(),
       y = 'SWE (mm)') +
  theme_bw(base_size = 20) +
  theme(
    #legend.position = c(0.89, 0.87),
    legend.position = 'bottom',
    legend.text = element_text(size = 10),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
    #plot.margin = margin(t = 10, r = 20, b = 10, l = 10)
  )

swe

p <- both %>%
  group_by(elev_source) %>%
  mutate(cumP = cumsum(precip)) %>% 
  ggplot(.,
       aes(x=date, y = cumP, 
           color = elev_source,
           linetype=elev_source)) +
   geom_line(size=1, alpha=0.8) +  
    scale_color_manual(values = c("Gridded - High" = "#0047AB", 
                                  "Gridded - Low" = "#0047AB", 
                                  "Station - High" = "orange",
                                  "Station - Low" = "orange")) +
  scale_linetype_manual(values = c("Gridded - High" = "dashed", 
                                  "Gridded - Low" = "solid", 
                                  "Station - High" = "dashed",
                                  "Station - Low" = "solid")) + 
  scale_x_date(date_labels = "%b") +
  #scale_y_log10() +
  labs(x = element_blank(),
       y = 'Precip (mm)') +
  theme_bw(base_size = 20) +
  theme(
    legend.position = 'bottom',
    #legend.text = element_text(size = 12),
    #legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
    #plot.margin = margin(t = 10, r = 20, b = 10, l = 10)
  )

p

legend <- cowplot::get_plot_component(tmean, "guide-box", return_all = TRUE)[[3]]

tmean_noleg <- tmean + theme(legend.position = "none")
p_noleg <- p + theme(legend.position = "none")
swe_noleg <- swe + theme(legend.position = "none")

combined <- plot_grid(tmean_noleg, p_noleg, swe_noleg, nrow = 1)
final_plot <- plot_grid(combined, legend, ncol = 1, rel_heights = c(1, 0.1))
final_plot

ggsave('./figures/figure7.png',
       width = 10,
       height = 5,
       dpi=800,
       bg='white')
